---
title: "HW 4"
output:
  word_document: default
  html_document: default
---

Problem 1
```{r}
pr8.4<- read.table("http://www.stat.umn.edu/~gary/book/fcdae.data/pr8.4",header=TRUE)
pr8.4$flow = as.factor(pr8.4$flow)
pr8.4$temp = as.factor(pr8.4$temp)
pr8.4$laser = as.factor(pr8.4$laser)
```


Assumption: We assume the standard three-way ANOVA model with interaction, with independent errors. We verify the assumptions of normality and constant variance with plots below:

```{r}
mod1 = aov(y ~ flow+temp+laser, data = pr8.4)
plot(mod1,which=c(1,2))
## Both the constant variance and QQ plot looks good 

anova(mod1)
```


Hypotheses:
Ho: All flow are the same 
Ha: At least one flow is different 

Ho: All temp are the same 
Ha: At least one temp is different 

Ho: All laser are the same 
Ha: At least one laser is different 


Conclusion: At the α = .05 level, we see that temp had a significance, therefore, we only have significant evidence of the temp treatment affecting delamination. 


We can look at the interaction plot 

The graphical method indicates that there is an interaction between flow and laser, but our statistical method was insignificant. 
```{r}
with(pr8.4,interaction.plot(temp,flow,y))
with(pr8.4,interaction.plot(temp,laser,y))
with(pr8.4,interaction.plot(flow,laser,y))

```

2.
Assumption:  We assume the standard three-way ANOVA model with interaction, with independent errors. We verify the assumptions of normality and constant variance with plots below:

```{r}
mod9 = aov(y ~ flow*temp*laser, data = pr8.4)
plot(mod9,which=c(1,2))
```

Hypotheses:
Ho: All flow are the same 
Ha: At least one flow is different 

Ho: All temp are the same 
Ha: At least one temp is different 

Ho: All laser are the same 
Ha: At least one laser is different

Ho: All interactions are the same 
Ha: At least one interaction is different 

```{r}
anova(mod9)
```
We see that the temp is lower than 0.05 significance level, so we reject the null hypothesis. Therefore, we can say that we have strong evidence that the temp has an effect on delamination. 



Problem 2 

```{r}
pr8.5 = read.table("http://www.stat.umn.edu/~gary/book/fcdae.data/pr8.5",header=TRUE)
pr8.5$shape = as.factor(pr8.5$shape)
pr8.5$trt = as.factor(pr8.5$trt)
```


Assumption: We assume the standard two-way ANOVA model with interaction, with independent errors. We verify the assumptions of normality and constant variance with plots:

```{r}
mod2 = lm(y ~ shape*trt, data = pr8.5)
plot(mod2,which=c(1,2))
```


Hypotheses: 
Ho: αshape = 0 vs H1: αshape =/ 0 
H0: βtrt = 0 vs H1: βtrt =/ 0 


```{r}
anova(mod2)
```

Conclusion: We see a significance in shape and treatment, therefore, we have significant evidence that both shape and treatment have an effect on resin 


```{r}
## We can visualize

##The interaction plot shows no interaction which is same as what we concluded. 
with(pr8.5,interaction.plot(shape,trt,y))
```


Problem 3 

```{r}
pr9.1 = read.table("http://www.stat.umn.edu/~gary/book/fcdae.data/pr9.1",header=TRUE)
pr9.1$method = as.factor(pr9.1$method)
pr9.1$time = as.factor(pr9.1$time)
pr9.1$type = as.factor(pr9.1$type)
```


Assumption: We assume the standard three-way ANOVA model with interaction, with independent errors. We verify the assumptions of normality and constant variance with plots:



```{r}
#The assumption of normality and variance looks violated, so we're going try transformations 
mod3 = lm(y ~ method+time+type, data = pr9.1)
plot(mod3,which=c(1,2))
```

Hypotheses: 
Ho: All method are the same 
Ha: At least one method is different 

Ho: All time are the same 
Ha: At least one time is different 

Ho: All type are the same 
Ha: At least one type is different 


```{r}
## log transform
mod4 = lm(log(y) ~ method+time+type, data = pr9.1)
par(mfrow = c(1,2))
plot(mod4,which=c(1,2))

## sqrt transform
tranmod1 = lm(sqrt(y) ~ method+time+type, data = pr9.1)
par(mfrow= c(1,2))
plot(tranmod1, c(1,2))

## Box Cox transform
library(car)
boxCox(tranmod1)
modd1 = lm(y^-2 ~ method+time+type, data = pr9.1)
plot(modd1, c(1,2))
```


It looks like the log transformation has the best transformation of the data because the variance bands look smoother 
and the QQ plot looks a little smoother than before. 
```{r}
anova(mod4)
```

Conclusion: At a 0.05 significance level, we reject method, time, and type. Therefore, we have significant evidence that method, time, and type all have an effect on the storage time. 


Problem 4 

```{r}
p10.3= read.table("http://www.stat.umn.edu/~gary/book/fcdae.data/pr10.3",header=TRUE)
p10.3$gum = as.factor(p10.3$gum)
p10.3$protein = as.factor(p10.3$protein)
```


Assumption: Because the data is unbalanced, we are going to use Type II SS as there were some missing values in the dataset

Hypotheses: Ho: αgump = 0 vs H1: αgum =/ 0 
H0: βprotein = 0 vs H1: βprotein =/ 0 



```{r}
Anova(lm(response ~ gum*protein, data = p10.3), type = 2)
```

We see gum p-value is lower than 0.05 significance level, therefore, we reject the null hypothesis. Meaning, that we have strong evidence that gum has an effect on the sensory rating. 

```{r}
## We see that gum2 had a significance, meaning that it differed the most in their sensory ratings
summary(lm(response ~ gum*protein, data = p10.3))
```


We can also look at the visualizations
From our graphical procedure, we do see an interaction, but from our statistical procedure we didn't. This might be because graphical procedures help you interpret the interaction effects, but statistical procedures actually determine whether the effect is statistically significant or not. 

```{r}
with(p10.3,interaction.plot(gum, protein,response))
```



