---
title: "HW 5"
output:
  word_document: default
  html_document: default
---

```{r}
ex13.2=read.table("http://www.stat.umn.edu/~gary/book/fcdae.data/ex13.2",header=TRUE)
ex13.2$batch = as.factor(ex13.2$batch)
```


Problem 1 

a) The variance component estimate is 5.423713

```{r}
ex13.2$batch = as.factor(ex13.2$batch)
ex13.2$temp = as.factor(ex13.2$temp)
ex13.2$pH = as.factor(ex13.2$pH)
ex13.2$sugar = as.factor(ex13.2$sugar)
```

```{r}
library(MASS)
mod1 = aov(y ~ batch+temp + pH + sugar, data = ex13.2)
summary(mod1)
plot(mod1)

boxcox(mod1, lambda = seq(0,3))

mod11 = (aov(y^1.5~batch + temp + pH + sugar, data = ex13.2))
summary(mod11)
plot(mod11)

## We can see that by doing the Box-Cox, our residuals are much better, and we meet the normalality assumption. 


(batch_var = (580 - 227)/12)
(sqrt(batch_var))
```

b)  batch variance = 29.43; batch sd = 5.425  

```{r}
library(lme4)
mod2 = lmer(y^1.5~ temp + pH + sugar + (1|batch), data = ex13.2)
summary(mod2)
```


c) We see that in part a and b, they produced similar results of 1.35. This is because the data is balanced, therefore the estimates align. 


d) ICC = 0.114719
```{r}
ICC = 29.43/(29.43+227.11)
ICC
```
e) ICC is the correlation of observations within a treatment group, therefore, because we see a low value, this might mean that there is not a similarity between observations within the same treatment group 


f)
Hypotheses: 
Ho:  σ2α = 0 vs. H1 : σ2α > 0
Ha: σ2α > 0


Test statistic/p-value

We see a p-value greater than 0.05, therefore, we fail to reject the null hypothesis and conclude that there is not significant evidence of nonzero variability between batches. 
```{r}
library(RLRsim)
exactRLRT(mod2)
```
g) We look at the sig01, and see that 0 is included in the interval, therefore, we fail to reject, which is the same as the results that we got in part f
```{r}
confint(mod2)
```

h) None of the factors nor combination of factors were significant that led to effecting the graininess 





Problem 2 

a) We see that the data is balanced
```{r}
data1 = read.csv("thermal.impedance.csv")
data1$inspector = as.factor(data1$inspector)
data1$test = as.factor(data1$test)
data1$ï..part = as.factor(data1$ï..part)

summary(data1)
```



```{r}
model.aov = aov(impedance ~ ï..part + inspector, data = data1)
summary(model.aov)

# check assumptions
## we see that the assumptions are pretty much met for normality and constant variance
par(mfrow=c(2,2))
plot(model.aov)
par(mfrow=c(2,2))
```


b) 
```{r}
mod3 = aov(impedance ~ ï..part + inspector + ï..part:inspector, data = data1)
summary(mod3)
```
```{r}
## variance component estimate 


## EMS(AB)
(2.6-0.5)/9

# EMS(A)
(438.8 - 0.5 - (9*0.2333333))/(9*3)

# EMS(B)
(19.9 - 0.5 -(9*0.2333333))/(9*6)
```

c) 0.6993 - ï..part:inspector variance
48.4639 - ï..part variance
0.5749 - inspector variance
```{r}
model.new = lmer(impedance ~ test +(1|ï..part) + (1|inspector) + (1|ï..part:inspector), data = data1) 
summary(model.new)
```

d) From the results, we see that all the values are less than 0.05, so it's statistically significant.
We conclude there is significant evidence of nonzero variability between part and inspector. 

H0 : σ2α = 0
H1 : σ2α > 0


```{r}

no.part = lmer(impedance ~ test + (1|inspector) + (1|ï..part:inspector), data = data1)
part = lmer(impedance ~ test + (1|ï..part), data = data1) 
exactRLRT(part,model.new, no.part )


no.inspector = lmer(impedance ~ test + (1|ï..part) + (1|inspector:ï..part), data = data1)
inspector = lmer(impedance ~ test+ (1|inspector), data = data1) 
exactRLRT(inspector, model.new, no.inspector)


no.part.by.inspector = lmer(impedance ~ test +(1|ï..part) + (1|inspector), data = data1)
part.by.inspector = lmer(impedance ~ test+(1|ï..part:inspector), data = data1) 
exactRLRT(part.by.inspector, model.new, no.part.by.inspector)
```


e) We see that sig01, sig02, sig03 do not contain 0, so we reject the null hypothesis, which is the same result that we got in part d
```{r}
confint(model.new)
```

Bonus













