---
title: "Homework 6"
output:
  word_document: default
  html_document: default
---

```{r}
data = read.csv("CashOffer.csv")
library(car)
```

Problem 1 

a)
We see that from the results, the p-values aren't significant. Therefore, the treatment doesn't affect the covariate 
We see that the sales volume is the covariate and it is statistically significant at pvalue = 4.11e-05

```{r}
data$age = as.factor(data$age)
data$gender = as.factor(data$gender)

mod5 = lm(offer ~ sales.volume, data = data) 
summary(mod5); anova(mod5)

mod6 = lm(sales.volume ~ age, data = data)
summary(mod6); anova(mod6) 

mod7 = lm(sales.volume ~ gender, data = data) 
summary(mod7); anova(mod7) 

```
Because the interaction is not significant, we use the parallel lines model 
```{r}
mod2 = aov(offer~ sales.volume*age*gender, data = data)
summary(mod2)
anova(mod2)
```
Now we're going to check the assumptions 

```{r}
## the assumptions look pretty good, but not great, so we're going to try BoxCox 
modex1 = lm(offer~sales.volume + age + gender, data = data)
plot(modex1, c(1,2))
```


```{r}
## from the BoxCox, the maximizing power transformation is about 1.5, which is pretty close to 1, so it’s unlikely to offer any improvement, but we can try looking at it
library(car)
boxCox(modex1)


## we see that using BoxCox, we saw a little improvement, but not a significant one.
example = lm(offer^1.5~sales.volume + age + gender, data = data)
plot(example, c(1,2))
```



```{r}
anova(modex1)
```

We see that age and gender was significant at the 0.05 significance level, therefore, both age and gender are important predictors of the cash offer. 


b) 
```{r}
final = (lm(offer~sales.volume + age+gender, data = data))
anova(final)
summary(final)
```

yijk(hat) = 18.16695 + 1.07393sales.volume  + 5.19676 (AgeMiddle) - 0.63062(AgeYoung) + 0.41383(GenderMale)


c) The experimental design would be two- way anova
age: categorical variable with no linear relationship with covariate
gender: categorical variable with no linear relationship with covariate


d)
```{r}
#levels(data$age) = list("Young" = 1, "Middle" =2, "Elderly" =3)
```

```{r}
plot(offer~sales.volume, col =age, pch =19, data = data)

abline(a = 18.16695, b =1.07393, col = "pink") 
abline(a = 18.16695 + 5.19676, b =1.07393, col = "yellow") 
abline(a = 18.16695 + 5.19676 +  0.41383, b =1.07393, col = "blue") 

abline(a = 18.16695 - 0.63262 + 0.41383 , b =1.07393, col = "green")
abline(a = 18.16695 - 0.63262 , b =1.07393, col = "red")
abline(a = 18.16695 + 0.41383, b = 1.07393, col = "black")

```

e) 
Young	Male :22.24388
Middle Female : 27.65943
Elderly Male : 22.87650
Young Femae : 21.83005
Middle Male : 28.07326
Elderly Female : 22.46267
```{r}
newdata = data.frame("sales.volume" = rep(4,6), "age" = c("Young", "Middle", "Elderly"), "gender" = c("Male", "Female"))
newdata

predict(modex1, newdata = newdata)
```



f) There are statistical differences among age with respect to the average cash offer because the p-values among age is less than 0.05. Therefore, Middle-Elderly, Young-Elderly, and Young-Middle are different 

```{r}
TukeyHSD(aov(final))
```

g) Young female receives the smallest cash offer on average
Middle Male receives the largest cash offer on average


h) Including the covariate reduced the MSE from 2.39 to 0.274, so it was decreased by 2.116



Problem 2 

```{r}
pr17.5 = read.table("http://www.stat.umn.edu/~gary/book/fcdae.data/pr17.5",header=TRUE)
```


```{r}
pr17.5$soil.type = as.factor(pr17.5$soil.type) 
pr17.5$distance = as.factor(pr17.5$distance)
```

```{r}
model = lm(merc.level ~ carb.frac, data = pr17.5)
Anova(model, type = 2)
summary(model)

model1 = lm(carb.frac ~ soil.type, data = pr17.5)
Anova(model1, type = 2) 
summary(model1) 

model2 = lm(carb.frac ~ distance, data = pr17.5) 
Anova(model2, type = 2) 
summary(model2) 

model3 = lm(merc.level ~ carb.frac*soil.type*distance, data = pr17.5)
Anova(model3, type=2)
summary(model3)
```

We see that there is significant linear relationship between soil type and the covariate

Because soil type effects the covariate, so we need to adjust the covariate value for each observation by the mean of the corresponding treatment.

```{r}
pr17.5$adjfeel = resid(lm(carb.frac ~ soil.type, data = pr17.5))

anova(lm(merc.level ~ adjfeel*soil.type*distance, data = pr17.5)) 
```
Because we see an interaction, we're going to use the separate lines model 


```{r}
# we see a violation for the constance variance, so we're going to try BoxCox

model4 = lm(merc.level ~ adjfeel * soil.type*distance, data = pr17.5)
plot(model4, c(1,2))

bc2 = boxCox(model4)
bc2$x[which.max(bc2$y)]


model5 = lm(merc.level^1.5 ~ adjfeel * soil.type*distance, data = pr17.5)
plot(model5, c(1,2))

model6 = lm(log(merc.level)~adjfeel*soil.type*distance,data=pr17.5)
plot(model6, c(1,2))
```


```{r}
model7 = lm(log(merc.level)~adjfeel*soil.type*distance,data=pr17.5)
Anova(model7, type = 2)
summary(model7)
```

```{r}
newdata3 = data.frame("adjfeel" = rep(mean(pr17.5$adjfeel),4), "soil.type"= as.factor(c(1,2)), "distance" = as.factor(c(1,2)))
predict(model7, newdata = newdata3)
```

yijk(hat) = -4.30582 +  28.86672(adjfeel) + 1.30767(soil.type2) - 0.17039(distance) - 20.84113(adjfeel:soil.type2) - 245.83309I2(adjfeel:distance2) + 0.08222(soiltype2:distance2) + 250.15026(adjfeel:soiltype2:distance2)


```{r}
plot(pr17.5$adjfeel, log(pr17.5$merc.level), col = pr17.5$soil.type, data = pr17.5)
abline(a = -4.30582 , b = 28.86672, col = "red")
abline(a =-4.30582+1.30767, b = 28.86672 -20.84113, col = "yellow") 
abline(a = -4.30582-0.17039 ,b = 28.86672 -245.83309  , col = "green")
abline(a = -4.30582+1.30767-0.17039, b = 28.86672 -20.84113 -245.83309 + 0.08222 +250.15026 , col = "blue")
```
From the graph, there is statistical evidence that both soil type and the distance can be the effect on the incinerator effect on soil mercury 


Bonus Problem 

```{r}
horse = read.csv("HorsePrices.csv")
horse$Sex = as.factor(horse$Sex)
```

a) Price = 16505 + 17225(sexM)

```{r}
# Because our pvalue is less than 0.05 significance level, sex has a significant effect on price 
modb = lm(Price ~ Sex, data = horse) 
summary(modb)
anova(modb)
```


```{r}
## the assumptions look good 
plot(modb, c(1,2))
```


b) 
```{r} 
horse$Sex = as.factor(horse$Sex) 

# We see that the two covariates are significant 
summary(lm(Price ~ Age, data = horse)) 
summary(lm(Price ~ Height, data = horse))


# We see that treatment doesn't effect covariate
equation1 = lm(Age ~ Sex, data = horse)
summary(equation1)
equation2 = lm(Height ~ Sex, data = horse)
summary(equation2)


## now see if there is a treatment by covariate interaction 
modb2 = lm(Price ~ Age*Height* Sex, data = horse) 
summary(modb2)


## we see that there is no significant interactions between treatment and covariate, therefore we use the parallel model 

equation3 = (lm(Price~ Age + Height+ Sex, data= horse))
anova(equation3)
summary(equation3)
```
Yij(hat) =  -136941.4 - 1329.2Age + 10067.9Height + 12572.6Sexm
If the gender is male, then we're going to see price increased by 12572.6


c) We saw a reduction in MSE from model b from 154906927 to 112360611


d) & e) 31869.03 = average price of male horses 
19296.45 = average price of female horses 
```{r}
newdata2 = data.frame("Age" = rep(mean(horse$Age),2), "Height" = rep(mean(horse$Height),2), "Sex" = as.factor(c("m", "f")))
newdata2
predict(equation3, newdata = newdata2)
```


f) 
```{r}
plot(horse$Age, horse$Price, data = horse, col = horse$Sex) 
plot(horse$Height, horse$Price, data = horse, col = horse$Sex)
```

```{r}
plot(horse$Age, horse$Price, col = horse$Sex, data = horse) 
abline(a = -136941.4+10067.9*(mean(horse$Height)), b = -1329.2, col = "red") 
abline(a = -136941.4  + 12572.6+ 10067.9*(mean(horse$Height)), b = -1329.2, col = "green") 
```


