---
title: "hw8"
output:
  word_document: default
  html_document: default
---

```{r}
library(MPV)
data(p13.5,package="MPV")
attach(p13.5)
```

Problem 1 


a) Model 2 is a better predictor of the vehicle purchase because there is less residual deviance in mod2 than mod1. 

```{r}

p13.5$y = as.factor(p13.5$y)
mod1 = glm(y ~ x1, family = "binomial", data = p13.5)
summary(mod1)

mod2 = glm(y~x2, family = "binomial", data = p13.5)
summary(mod2)

anova(mod1, mod2, test = "Chisq")
```

b) We see that mod2 is better with vehicle age as the predictor because our p-value is not significant
```{r}
mod3 = glm(y ~ x1 + x2, family = "binomial", data = p13.5)
summary(mod3)

anova(mod2, mod3, test = "Chisq")
```

c) The probability of purchasing a new vehicle increases with increasing vehicle age 
B1:  0.555968 
for every 1 unit increase in vehicle age, the log odds increases by 0.555968
```{r}
coef(mod2)
```

d) Adding the interaction term is better than the previous model beccause our pvalue is less than 0.05
```{r}
mod4 = glm(y ~ x1 * x2, family = "binomial", data = p13.5)
summary(mod4)

anova(mod2, mod4, test = "Chisq")
```

e) null_deviance =  27.72588
residual_deviance =  16.5508
```{r}
sat = glm(y~as.factor(1:length(y)), data = p13.5, family = "binomial")

residual_deviance = 2*(logLik(sat) - logLik(mod4))


null = glm(y~1, data =p13.5, family = "binomial")
logLik(null)

null_deviance = 2*(0--13.86294)
```


f) AIC = 24.5508
```{r}
residual_deviance+2*4
```

g) The estimated probability is 0.9886166 
```{r}
predict(mod4, data.frame(x1 = 45000, x2 = 5), type = "response")
```

Problem 2 
```{r}
data(p13.4,package="MPV")
attach(p13.4)
```

a) With every 1 unit increase in x, the log odds increases by 0.1357274
We also see that when there's more discount on the coupons, people redeem more coupons
```{r}
p13.4$prop = p13.4$r/p13.4$n
p13.4$nr = n-r

mod5 = glm(cbind(r,nr) ~ x, data = p13.4, family = "binomial")
summary(mod5)
coef(mod5)
```

b) 15.35986
```{r}
log(0.5/0.5)
discount1 = -( -2.0847543 - log(0.5/0.5)) /  0.1357274 
```


c) 25.57368
```{r}
log(0.8/0.2) 
discount2 = -(-2.0847543 - log(0.8/0.2)) / 0.1357274 
```



Problem 3 

```{r}
library(alr4)
data(Downer,package="alr4")
attach(Downer)
```




a) The coefficient for inflamat is negative. This means that the probability of outcome of animals decreases when the inflamat is present. 
```{r}

Downer$outcome = as.factor(Downer$outcome)
Downer$inflamat = as.factor(Downer$inflamat)

mod6 = glm(outcome ~ inflamat, family = "binomial", data = Downer)
summary(mod6)

coef(mod6)
```

b) The probability that a cow with inflammation will survive is 30.61%
0.3061224
```{r}
prob_yes = predict(mod6, data.frame(inflamat="yes"), type = "response", data = p13.4)
```

c) The probability that a cow without inflammation will survive is 44.73%

0.4473684 
```{r}
prob_no = predict(mod6, data.frame(inflamat="no"), type = "response", data = p13.4)

```


d) The odds ratio was -0.6070, so having inflamat is more likely to survive 0.54 times than not having inflamat
```{r}
oddsratio = exp(-0.6070)
```

e) Agrees with part d
```{r}
table(Downer$inflamat, Downer$outcome)

or = (30/68) / (17/21)
oddsratio
```


f) 
```{r}
log(17/21)
```

g) Outcome = 0.5284900 + (-0.6478436)inflamatyes + (-0.0002848)ck
```{r}
mod7 = glm(outcome ~ inflamat + ck, data = Downer, family = "binomial")
summary(mod7)
```






